---
title: "Milestone-2-rye"
author: "Erik Rye"
date: "2023-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caTools)
library(dplyr)
```

## Milestone 2 

### Linear Regressions

### Question 1: Can we predict the number of applicants a job post will receive given the number of views that it receives?

```{r}

# Set current working directory to where my data is
setwd('/home/erik/repos/gigaryte/INST737-Team-8/')
# read the main CSV into a dataframe
df <- read.csv('job_postings.csv', header=TRUE)

applications <- data.frame(views = df$views, applies = df$applies)
applications <- na.omit(applications)

# Seed the test/train sample so we have repeatable results
set.seed(0)

# Do a 70/30 training/testing split
sample <- sample.split(applications$views, SplitRatio = 0.7)
train  <- subset(applications, sample == TRUE)
test   <- subset(applications, sample == FALSE)

# Plot the training data
plot(train$views, train$applies, 
     xlab = "Number of Views", 
     ylab = "Number of Applications", 
     pch=16,
     main = "Training Data w/Linear Regression") 

# Linear regression model for predicting applies based on views trained on training data
applies.lm <- lm(applies~views, data = train)
applies.lm

# Plot the linear regression on same axes
abline(applies.lm, col="red", lwd=3)

# Create a function to run on test data
view.fit <- function(x) applies.lm$coefficients[1] + applies.lm$coefficients[2] * x

# Add the fit and residuals to the dataframe
train <- data.frame(train, fit = fitted(applies.lm), resid= resid(applies.lm))

# Plot the Q-Q plot
qqnorm(resid(applies.lm))
qqline(resid(applies.lm))

########## Predictions and compare with test data ##########

# Predict the number of applications given the number of views from our test data
# using our linear regression model
prediction.frame <- data.frame(views = test$views)
pc <- predict(applies.lm, int='c', newdata = prediction.frame)
pp <- predict(applies.lm, int='p', newdata = prediction.frame)
plot(train$views, train$applies, 
     xlab = "Number of Views", 
     ylab = "Number of Applications", 
     main = "Test Data with Prediction and Confidence Bands",
     pch=16)
matlines(prediction.frame, pc)
matlines(prediction.frame, pp)

# Get the predictions for the number of applications based on the number
# of views in our test set
applies_prediction <- predict(applies.lm, newdata=prediction.frame)

# See how well it correlates with the actual test applications
cor(applies_prediction, test$applies)

# Calculate residuals of test data 
test_residuals <- test$applies - applies_prediction
mse <- mean(test_residuals^2)
mse

```

##  Can we predict the number of applicants based on the job's salary information, including both salary range and whether the salary is listed vs unlisted?

```{r}
# Set current working directory to where my data is
setwd('/home/erik/repos/gigaryte/INST737-Team-8/')
# read the main CSV into a dataframe
df <- read.csv('job_postings.csv', header=TRUE)

  applications <- data.frame(applies = df$applies, salary = df$salary)
applications <- na.omit(applications)

# Seed the test/train sample so we have repeatable results
set.seed(0)

# Do a 70/30 training/testing split
sample <- sample.split(applications$views, SplitRatio = 0.7)
train  <- subset(applications, sample == TRUE)
test   <- subset(applications, sample == FALSE)

# Plot the training data
plot(train$views, train$applies, 
     xlab = "Number of Views", 
     ylab = "Number of Applications", 
     pch=16,
     main = "Training Data w/Linear Regression") 

# Linear regression model for predicting applies based on views trained on training data
applies.lm <- lm(applies~views, data = train)
applies.lm

# Plot the linear regression on same axes
abline(applies.lm, col="red", lwd=3)

# Create a function to run on test data
view.fit <- function(x) applies.lm$coefficients[1] + applies.lm$coefficients[2] * x

# Add the fit and residuals to the dataframe
train <- data.frame(train, fit = fitted(applies.lm), resid= resid(applies.lm))

# Plot the Q-Q plot
qqnorm(resid(applies.lm))
qqline(resid(applies.lm))

########## Predictions and compare with test data ##########

# Predict the number of applications given the number of views from our test data
# using our linear regression model
prediction.frame <- data.frame(views = test$views)
pc <- predict(applies.lm, int='c', newdata = prediction.frame)
pp <- predict(applies.lm, int='p', newdata = prediction.frame)
plot(train$views, train$applies, 
     xlab = "Number of Views", 
     ylab = "Number of Applications", 
     main = "Test Data with Prediction and Confidence Bands",
     pch=16)
matlines(prediction.frame, pc)
matlines(prediction.frame, pp)

# Get the predictions for the number of applications based on the number
# of views in our test set
applies_prediction <- predict(applies.lm, newdata=prediction.frame)
applies_residuals <- resid(applies.lm, newdata=)

# See how well it correlates with the actual test applications
cor(applies_prediction, test$applies)

# Calculate residuals of test data 
test_residuals <- test$applies - applies_prediction
mse <- mean(test_residuals^2)
mse
```

###  Can we predict a job's salary based on the industry the job is from and where the job is located?

```{r}

```


###  Can we predict the number of views a job advertisement will receive based on the size of the company posting the ad?

```{r}
# Set current working directory to where my data is
setwd('/home/erik/repos/gigaryte/INST737-Team-8/')
# read the main CSV into a dataframe
df <- read.csv('job_postings.csv', header=TRUE)

employee_info <- read.csv('company_details/employee_counts_clean.csv', header=TRUE)

df <- merge(df, employee_info, by.x = "company_id", by.y = "company_id", all.x = FALSE)

views <- data.frame(views = df$views, employees = df$employee_count)
views <- na.omit(views)

# Seed the test/train sample so we have repeatable results
set.seed(0)

# Do a 70/30 training/testing split
sample <- sample.split(views$views, SplitRatio = 0.7)
train  <- subset(views, sample == TRUE)
test   <- subset(views, sample == FALSE)

# Plot the training data
plot(train$employees, train$views, 
     xlab = "Number of Employees", 
     ylab = "Number of Views", 
     pch=16,
     main = "Training Data w/Linear Regression") 

# Linear regression model for predicting applies based on views trained on training data
views.lm <- lm(views~employees, data = train)
views.lm

# Plot the linear regression on same axes
abline(views.lm, col="red", lwd=3)

# Create a function to run on test data
view.fit <- function(x) views.lm$coefficients[1] + views.lm$coefficients[2] * x

# Add the fit and residuals to the dataframe
train <- data.frame(train, fit = fitted(views.lm), resid= resid(views.lm))

# Plot the Q-Q plot
qqnorm(resid(views.lm))
qqline(resid(views.lm))

########## Predictions and compare with test data ##########

# Predict the number of views given the number of employees from our test data
# using our linear regression model
prediction.frame <- data.frame(views = test$views, employees = test$employees)
pc <- predict(views.lm, int='c', newdata = prediction.frame)
pp <- predict(views.lm, int='p', newdata = prediction.frame)
plot(train$employees, train$views, 
     xlab = "Number of Employees", 
     ylab = "Number of Views", 
     main = "Test Data with Prediction and Confidence Bands",
     pch=16)
matlines(prediction.frame, pc)
matlines(prediction.frame, pp)

# Get the predictions for the number of views based on the number
# of employees in our test set
views_prediction <- predict(views.lm, newdata=prediction.frame)

# See how well it correlates with the actual test views
cor(views_prediction, test$views)

# Calculate residuals of test data 
test_residuals <- test$views - views_prediction
mse <- mean(test_residuals^2)
mse
```



