---
title: "Feature Selection - Sharvil"
author: "Sharvil"
date: "2023-11-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Question 5. Feature Selection
Select three models from Milestones 2 or 3 (linear/logistic regression, Naïve Bayes, Decision Trees, SVM or Neural Networks) and apply three different types of feature selection: filter, wrapper or embedding, one per model. Report results focusing on improvements in performance (or lack of) due to feature selection.

```{r}
#Importing the libraries
library(dplyr)
library(tidyverse)
library(MASS)
library(ggplot2)
```

```{r}


#Importing the data. (Please select job_postings.csv file after running the read.csv function)
dataset <- read.csv(file.choose(), header = TRUE)

#Removing Scientific notations
options(scipen = 999)
```



#1 FS-SFS
In Milestone 2, for Multivariate Linear Regression, we had created a model, where the Dependent variable was the number of applicants & the independent variables were the number of views, salary, the experience level, the work type & the state.

The is the code for the model we used in Milestone 2. We have adjusted the variable names for convenience.
```{r}
attach(dataset)
dataset_FS1 <- data.frame(applies,views,clean_salary,formatted_experience_level,formatted_work_type,state)


#Cleaning up the data
dataset_FS1$formatted_experience_level[dataset_FS1$formatted_experience_level==""] <- NA
dataset_FS1$state[dataset_FS1$state==""] <- NA
dataset_FS1$clean_salary <- as.integer(dataset_FS1$clean_salary)
dataset_FS1  <- drop_na(dataset_FS1 )
#There should be 2334 observations of 6 variables in dataset_FS1
```

In the code below, we carry out the Feature Selection:
```{r}
#Step 1: Define base intercept only model (no variables). This model does not have any features. Just intercept
base.mod <- lm(applies ~ 1, data=dataset_FS1)

#Step 2: Full model with all predictors
all.mod <- lm(applies ~., data=dataset_FS1)

#Step 3: perform step-wise algorithm. direction = Forward
stepMod <- step(base.mod, scope=list(lower=base.mod, upper = all.mod), direction = "forward", trace = 0, steps = 1000)

#step() function in R searches for the best possible model by iteratively selecting features to arrive at a model with the lowest possible AIC


#Step 4: Get the shortlisted variable.
shortlistedVars <- names(unlist(stepMod[[1]]))
shortlistedVars <- shortlistedVars[!shortlistedVars %in% "(Intercept)"]
#Remove intercept

#Show
print(shortlistedVars)
#Basically, when we print this, whatever results we get in the console, those are actually the variables that give us the best model


#Model
summary(stepMod)
summary(all.mod)

```
With forward approach, we get an Adjusted R-squared value of around 0.7581


```{r}
#This example is for backward approach. We start with a model with all the features.
stepMod_backward <- step(all.mod, direction = "backward", trace = 0, steps = 1000)

#Get the shortlisted variable.

shortlistedVars2 <- names(unlist(stepMod_backward[[1]]))
shortlistedVars2 <- shortlistedVars2[!shortlistedVars2 %in% "(Intercept)"]
#Remove intercept

#Show
print(shortlistedVars2)
#Basically, when we print this, whatever results we get in the console, those are actually the variables that give us the best model


#Model
summary(stepMod_backward)
```
Both Backward & Forward approaches produce the same shortlisted variables, and the summary of both of them gives an R-squared value of around 0.7581
The R-squared value for the multivariate regression model was around 0.7592. With the Feature Selection method, it decreases very slightly. 


#Neural Networks:
This was the code which gave the best result for Neural Networks:

```{r}

#Loading the libraries
library(neuralnet)
library(sigmoid)
#Importing the data & Creating the specific dataframe:
jobs <- read.csv('job_postings.csv', header=TRUE)
employees <- read.csv('employee_counts_clean.csv', header=TRUE)
# Combine the two csvs on company_id
job_employees <- merge(jobs, employees, by.x = "company_id", by.y = "company_id", all.x = TRUE)
# Creating a data frame out of just the dependent variables
df <- data.frame(views=job_employees$views, 
                   salary=job_employees$clean_salary,
                   employees=job_employees$employee_count, 
                   followers=job_employees$follower_count, applies=job_employees$applies 
                 )
library(tidyverse)
df <- drop_na(df)

#Normalizing the continuous variables
normalize <- function(x) {
  return((x-min(x))/(max(x) - min(x)))
}
#Using lapply function to apply function to whole data frame
normalized_data <- as.data.frame(lapply(df,normalize))

#Creating the Training & Testing Datasets
applicants_train <- normalized_data[1:2500,]
applicants_test <- normalized_data[2501:3251,]
set.seed(123)

#Code for best Neural Network Model
applicants_model_main <- neuralnet(formula = applies ~ views + salary + employees + followers, data = applicants_train, hidden = c(3,3,2,2))

#Plotting the Neural Network Model
plot(applicants_model_main)

#Predicting, using the testing dataset
model_main_results <- neuralnet::compute(applicants_model_main, applicants_test[1:4])
predicted_applicants_model_main <- model_main_results$net.result
cor(predicted_applicants_model_main,applicants_test$applies)
```
The Pearson correlation Cofficient value should be around 0.9 or 0.89


Now, creating the cor plot

```{r}
corr_variables <- cor(applicants_train)

#Building correplot to visualize the correlation matrix. Show numbers (not colors);
#Legend value range
library(corrplot)
corrplot(cor(applicants_train), method="number", is.corr = FALSE)

```
As we can see from the plot, only views variable has a strong correlation with the applies column. We can eliminate all the other ones. (salary, employees,and followers)


Let's create the Neural Network model using only the views columns.
```{r}

library(neuralnet)
#Only using views column as the IV

model2 = update(applicants_model_main, ~.-salary-employees-followers)
#Plotting the Neural Network Model
plot(model2)
model2_main_results <- neuralnet::compute(model2, applicants_test[1])
#applicants_test[1] because the 1st column in the dataframe is 'views' column.
predicted_applicants_model_main <- model2_main_results$net.result
cor(predicted_applicants_model_main,applicants_test$applies)
```

The correlation Coefficient value comes around 0.89, or sometimes 0.88

The code below is for the neural network model, with feature selection, WITHOUT any hidden layers or nodes:
```{r}
applicants_model_main_2 <- neuralnet(formula = applies ~ views + salary + employees + followers, data = applicants_train)
model3 = update(applicants_model_main_2, ~.-salary-employees-followers)
#Plotting the Neural Network Model
plot(model3)
model3_main_results <- neuralnet::compute(model3, applicants_test[1])
#applicants_test[1] because the 1st column in the dataframe is 'views' column.
predicted_applicants_model_main_2 <- model3_main_results$net.result
cor(predicted_applicants_model_main_2,applicants_test$applies)
```

The correlation coefficient value even without any hidden layers or nodes, is more or less the same in this case too.




#3. Decision Trees


```{r}
#Importing the Libraries
library(caTools)
library(C50)
library(gmodels)
library(tidyverse)

# Add the number of employees
job_employees <- merge(jobs, employees, by.x = "company_id", by.y = "company_id", all.x = TRUE)


# Filter for the columns we want, and remove "class_applies" unspecifieds for
# the purposes of prediction
filtered_df <- job_employees  %>%
      dplyr::select(class_views, class_applies, class_salary, formatted_work_type, remote_allowed, class_employees, class_followers) %>% 
        mutate(remote_allowed = ifelse(is.na(remote_allowed), 0, remote_allowed)) %>% 
          filter(class_applies != "Unspecified")


      filtered_df$class_views <- as.factor(filtered_df$class_views)
      filtered_df$class_applies <- as.factor(filtered_df$class_applies)
      filtered_df$class_salary <- as.factor(filtered_df$class_salary)
      filtered_df$class_employees <- as.factor(filtered_df$class_employees)
      filtered_df$class_followers<- as.factor(filtered_df$class_followers)
      filtered_df <- drop_na(filtered_df)
      # Seed the test/train sample so we have repeatable results
      set.seed(1)

     
     
      sample <- sample.split(filtered_df$formatted_work_type, SplitRatio = 0.7)
      train  <- subset(filtered_df, sample == TRUE)
      test   <- subset(filtered_df, sample == FALSE)

      applies_model <- C5.0(train[-2], train$class_applies)
      summary(applies_model)

      applies_predictions <- predict(applies_model, test)
              CrossTable(test$class_applies, applies_predictions,
                 prop.chisq = FALSE, prop.c = FALSE,
                 prop.r = FALSE,
                 dnn = c('actual applications', 'predicted applications'))
      
      

```


Now, we will create the confusion matrix:
```{r}
library(caret)
confusionMatrix(test$class_applies, applies_predictions, mode="everything")
#The Accuracy value should come around 0.72 or 0.73 when running the above code

```


Now, we will use Random Forest Feature Selection Embedded Method:

```{r}
library(randomForest)
#Creating the Random Forest Model:
random_forest_model <- randomForest(class_applies ~ ., data = train)

#Getting the value of measures for the most important variables
importance(random_forest_model)

```
The MeanDecreaseGini value of class_views is significantly higher compared to other variables. Hence, all other variables are discarded.

```{r}
#Creating new decision tree model, with only the class_views as the predictor

      applies_model_2 <- C5.0(train[1], train$class_applies)
      summary(applies_model_2)

      #Predicting the values in the testing dataset
      applies_predictions_2 <- predict(applies_model_2, test)
      
      #Creating the CrossTable
              CrossTable(test$class_applies, applies_predictions_2,
                 prop.chisq = FALSE, prop.c = FALSE,
                 prop.r = FALSE,
                 dnn = c('actual applications', 'predicted applications'))
              
#Getting the Overall Statistics value              
confusionMatrix(test$class_applies, applies_predictions_2, mode="everything")
```
The accuracy value after Feature Selection should be around 0.702.